---
title: "Notebook for project exploratory data analysis"
output:
  html_document:
    df_print: paged
  pdf_document: default
---

```{r}
library(ggplot2)
library(tidyverse)
library(reshape2)
library(caret)
library(gt)
```

```{r}
train <- read.csv('train.csv')
numeric.cols <- names(train[, sapply(train, is.numeric)])
cat.cols <- names(train)[!(names(train) %in%  numeric.cols)]
```

## Exploratory Data Analysis

We're dealing with some missing values in `Garage_Yr_Blt` variable. We'll have 
to deal with this somehow.

```{r}
train.na <- sapply(train, function(x) sum(is.na(x)))
train.na[train.na > 0]
```

Next, we'll just take a look at how `Sale_Price` is distributed and what it might
correlate it since that's the target variable.

```{r}
ggplot(data = train, aes(x = Sale_Price)) +
  geom_histogram(bins = 50)
```

```{r}
train$Sale_Price <- log1p(train$Sale_Price)
ggplot(data = train, aes(x = Sale_Price)) +
  geom_histogram(bins = 50)

```

There are a lot of numeric variables, so we'll split the columns in two lists
to better view the correlation matrices.

```{r}
# Get lower triangle of the correlation matrix
get_lower_tri<-function(cormat){
  cormat[upper.tri(cormat)] <- NA
  return(cormat)
}
# Get upper triangle of the correlation matrix
get_upper_tri <- function(cormat){
  cormat[lower.tri(cormat)]<- NA
  return(cormat)
}

# Cluster values in corr matrix
reorder_cormat <- function(cormat){
  # Use correlation between variables as distance
  dd <- as.dist((1-cormat)/2)
  hc <- hclust(dd)
  cormat <-cormat[hc$order, hc$order]
}

# Corr matrix gets hard to read with too many vars so cut down
cols1 <- c(numeric.cols[1:9], "Sale_Price")
cols2 <- c(numeric.cols[10:18], "Sale_Price")
cols3 <- c(numeric.cols[19:27], "Sale_Price")
cols4 <- c(numeric.cols[27:length(numeric.cols)])
```

```{r}
  
PlotCormat <- function(cols, reorder = FALSE) {
  cormat <- round(cor(train[, cols]), 3)
  
  if (reorder == TRUE) {
    cormat <- reorder_cormat(cormat)
  }
  upper_tri <- get_upper_tri(cormat)
  lower_tri <- get_lower_tri(cormat)
  cormat.melt <- melt(lower_tri, na.rm = TRUE)
  
  ggplot(data = cormat.melt, aes(x=Var1, y=Var2, fill=value)) + 
    geom_tile(color = "white") +
    geom_text(aes(x=Var1, y=Var2, label=value), color = "black", size = 4) +
    scale_fill_gradient2(low = "blue", high = "red", mid = "white", midpoint = 0,
                         limit = c(-1, 1), space = "Lab", name = "Correlation") +
    theme_minimal() +
    theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1),
          axis.title.x = element_blank(),
          axis.title.y = element_blank(),
          panel.grid.major = element_blank(),
          panel.border = element_blank(),
          legend.direction = "horizontal",
          legend.justification = c(1, 0),
          legend.position = c(0.6, 0.7))
}
```

```{r}
PlotCormat(cols1, reorder = TRUE)
```


```{r}
PlotCormat(cols2)
```

```{r}
PlotCormat(cols3)
```

```{r}
PlotCormat(cols4)
```


```{r}
ggplot(data = train, aes(x=Latitude, y = Longitude, color = Sale_Price)) +
  geom_point(alpha = 0.5)
```

```{r}
res <- as.data.frame(read.csv('results_all_2.csv'))
ggplot(res, aes(x=test_set, y=rmse, fill = model_type)) + 
  geom_col(position="dodge") +
  theme_minimal() +
  theme(panel.grid.major = element_blank(),
        panel.border = element_blank(),
        legend.direction = "horizontal",
        legend.justification = c(1, 0),
        legend.position = c(0.6, 0.7)) +
  scale_fill_brewer(palette="Set1") +
  scale_x_discrete(limits = 1:10) +
  theme(legend.position="bottom") + 
  coord_cartesian(ylim = c(0.105, .135)) +
  xlab("Test Set Number") + 
  ylab("RMSLE") + 
  ggtitle("Sample RMSLE Results for All Test Sets") +
  theme(plot.title = element_text(hjust = 0.5)) + 
  geom_segment(aes(x=0, xend=5.5, y=.125, yend=.125)) +
  geom_segment(aes(x=5.5, xend=10.5, y=.135, yend=.135))
  #geom_text(aes(label=rmse), size=2) +
  #coord_flip()

```

```{r}
gt(res) %>%
  tab_header(
    title = md("Model Test Results"),
    subtitle = "RMSLE Results and Run Times for Training and Predictions"
  )
```

