---
title: "Movielense Data: Recommender System"
date: "CS598, Spring 2021"
output:
  html_notebook:
    theme: readable
    toc: TRUE
    toc_float: TRUE
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
```

```{r, message=FALSE}
library(recommenderlab)
library(Matrix)
```

## Read in Data

```{r}
myurl = "https://liangfgithub.github.io/MovieData/"

ratings = read.csv(paste0(myurl, 'ratings.dat?raw=true'), 
                   sep = ':',
                   colClasses = c('integer', 'NULL'), 
                   header = FALSE)
colnames(ratings) = c('UserID', 'MovieID', 'Rating', 'Timestamp')
ratings$Timestamp = NULL
```


## Evaluation I

### Training Test Split

Create

* **train** data that contain about 80\% rows of `ratings.dat` 

* **test** data that contain about 20\% of rows of `ratings.dat`


```{r}
set.seed(100)
train.id = sample(nrow(ratings), floor(nrow(ratings)) * 0.8)
train = ratings[train.id, ]
head(train)

test = ratings[-train.id, ]
head(test)
```

### Recommender System

* First create a utility matrix stored as a sparse matrix. 

```{r}
i = paste0('u', train$UserID)
j = paste0('m', train$MovieID)
x = train$Rating
tmp = data.frame(i, j, x, stringsAsFactors = T)
Rmat = sparseMatrix(as.integer(tmp$i), as.integer(tmp$j), x = tmp$x)
rownames(Rmat) = levels(tmp$i)
colnames(Rmat) = levels(tmp$j)
Rmat = new('realRatingMatrix', data = Rmat)
```

`Rmat` is a 6040-by-3681 sparse matrix. Its rows correspond to the unique 6040 users in the training data, and columns correspond to the unique 3681 movies  in the training data (although the MovieIDs range from 1 to 3952). 

```{r}
range(train$MovieID)
length(unique(train$MovieID))
range(train$UserID)
length(unique(train$UserID))
dim(Rmat)
```


* Train a recommender system using R package [[recommenderlab](https://cran.r-project.org/web/packages/recommenderlab/vignettes/recommenderlab.pdf)]

```{r}
rec_UBCF = Recommender(Rmat, method = 'UBCF',
                  parameter = list(normalize = 'Z-score', 
                                   method = 'Cosine', 
                                   nn = 25))
#Summary of model parameters
rec_UBCF@model
```

* Make prediction on the test data

First let's predict all un-rated movies for the first three users in the training data. No information for users not in the training data; the coldstart issue. 

```{r}
# predict ratings for the first three users
recom = predict(rec_UBCF, Rmat[1:3], type = 'ratings')
```

We display the original and predicted ratings of the first 10 movies for the three users. Note that the final prediction could still have missing values. You can consider filling in the missing values with some constant. 

```{r}
as(recom, 'matrix')[, 1:10]
as(Rmat, 'matrix')[1:3, 1:10]
```

You can use the code blow to compute RMSE for the test data. Note that this evaluation **takes a long time**. 


```{r, eval=FALSE}
# This may take a long time
recom = predict(rec_UBCF, Rmat, type = 'ratings')  
rec_list = as(recom, 'list')  # each element are ratings of that user

(rec_list[['u1']]['m2355'])

test.pred = test
test.pred$rating = NA

# For all lines in test file, one by one
for (u in 1:nrow(test)){

    # Read userid and movieid from columns 2 and 3 of test data
    userid = paste0('u',as.character(test$UserID[u]))
    movieid = paste0('m',as.character(test$MovieID[u]))
    
    rating = rec_list[[userid]][movieid]
    # handle missing values; 2.5 might not be the ideal choice
    test.pred$rating[u] = ifelse(is.na(rating), 2.5, rating)
}

# Calculate RMSE
sqrt(mean((test$Rating - test.pred$rating)^2)) 
```

All algorithms available in `recommenderlab`. Check  [[recommenderlab](https://cran.r-project.org/web/packages/recommenderlab/vignettes/recommenderlab.pdf)] for technical details. 

```{r}
recommenderRegistry$get_entry_names()
```

## Evaluation II

The previous evaluation scheme takes a long time since it predicts all missing entries in a large sparse matrix. A computation-friendly scheme is to form the training and test datasets by spliting users: 

* data from training users  are used to build a recommendation system; 
* data from the test users are split into two parts;

* one part is provided to this recommendation system for us to obtain the prediction on the other part so we can evaluate the prediction accuracy. 

This evaluation scheme can be called by the built-in evaluation function in `recommenderlab`; check Chap 3 of [[recommenderlab](https://cran.r-project.org/web/packages/recommenderlab/vignettes/recommenderlab.pdf)].

